IGAS Project

### 2024-04-11 Sections/Introduction.qmd
generate md of IGAS
intro update

Varied training has been shown to influence learning in wide array of different tasks and domains, including categorization [@hahnEffectsCategoryDiversity2005; @maddoxStimulusRangeDiscontinuity2011; @posnerGenesisAbstractIdeas1968; @nosofskyModelguidedSearchOptimal2019], language learning [@jonesDensityDistinctivenessEarly2020; @perryLearnLocallyThink2010; @twomeyAllRightNoises2018; @wonnacottInputEffectsAcquisition2012], anagram completion [@goodeSuperiorityVariableRepeated2008], trajectory extrapolation [@fulvioTaskSpecificResponseStrategy2014], task switching [@sabahWhenLessMore2019], associative learning [@leeEvidentialDiversityIncreases2019], visual search [@georgeStimulusVariabilityTask2021; @gonzalezDiversityTrainingEnhances2011; @kelleyLearningAttendEffects2009], voice identity learning [@lavanEffectsHighVariability2019], simple motor learning [@braunMotorTaskVariation2009; @kerrSpecificVariedPractice1978; @rollerVariablePracticeLenses2001; @willeyLimitedGeneralizationVaried2018], sports training  [@greenPracticeVariabilityTransfer1995a; @northEffectConsistentVaried2019], and training on a complex video game [@seowTransferEffectsVaried2019].

Research on the effects of varied training typically compares a group of participants trained with minimal variability of examples against a group with greater variability [@czyzVariabilityPracticeInformation2021; @ravivHowVariabilityShapes2022].

@@

On occasion the Kerr and Booth design may be nested within a larger experimental design. One such study that used a movement timing task, wherein subjects had to move their hand from a starting location, to a target location, attempting to arrive at the target location at specific time following the onset of a cue [@wrisbergVariabilityPracticeHypothesis1987]. This study utilized 4 different constant groups, and 3 varied groups, with one of the constant groups training under conditions identical to the testing conditions, and which were not trained on by any of the varied groups, e.g. the design of Kerr and Booth. However, in this case the varied group did not outperform the constant group. A more recent study attempting a slightly more direct replication of the original Kerr & Booth study [@willeyLongtermMotorLearning2018], having subjects throw beanbags at a target, with the varied group training from positions (5 and 9 feet) on either side of the constant group (7 feet). However, this study diverged from the original in that the participants were adults; they faced away from the target and threw the beanbag backwards over their bodies; they alternated using their right and left hands every 6 trials; and underwent a relatively extreme amount of training (20 sessions with 60 practice trials each, spread out over 5-7 weeks). Like @wrisbergVariabilityPracticeHypothesis1987, this study did not find a varied advantage from the constant training position, though the varied group did perform better at distances novel to both groups.

>>>>

Varied training has been shown to influence learning in wide array of different tasks and domains, including categorization [@hahnEffectsCategoryDiversity2005; @maddoxStimulusRangeDiscontinuity2011; @posnerGenesisAbstractIdeas1968; @nosofskyModelguidedSearchOptimal2019], language learning [@jonesDensityDistinctivenessEarly2020; @perryLearnLocallyThink2010; @twomeyAllRightNoises2018; @wonnacottInputEffectsAcquisition2012], anagram completion [@goodeSuperiorityVariableRepeated2008], trajectory extrapolation [@fulvioTaskSpecificResponseStrategy2014], task switching [@sabahWhenLessMore2019], associative learning [@leeEvidentialDiversityIncreases2019], visual search [@georgeStimulusVariabilityTask2021; @gonzalezDiversityTrainingEnhances2011; @kelleyLearningAttendEffects2009], voice identity learning [@lavanEffectsHighVariability2019], simple motor learning [@braunMotorTaskVariation2009; @kerrSpecificVariedPractice1978; @rollerVariablePracticeLenses2001; @willeyLimitedGeneralizationVaried2018], sports training  [@greenPracticeVariabilityTransfer1995a; @northEffectConsistentVaried2019], and training on a complex video game [@seowTransferEffectsVaried2019]. See @czyzVariabilityPracticeInformation2021 or @ravivHowVariabilityShapes2022 for more detailed reviews.

Research on the effects of varied training typically manipulates variability in one of two ways. In the first approach, a high variability group is exposed to a greater number of unique instances during training, while a low variability group receives fewer unique instances with more repetitions. Alternatively, both groups may receive the same number of unique instances, but the high variability group's instances are more widely distributed or spread out in the relevant psychological space, while the low variability group's instances are clustered more tightly together. Researchers then compare the training groups in terms of their performance during the training phase, as well as their generalization performance during a testing phase. Researchers will usually compare the performance of the two groups during both the training phase and a subsequent testing phase. The primary theoretical interest is often to assess the influence of training variability on generalization to novel testing items or conditions. However, the test may also include some or all of the items that were used during the training stage, allowing for an assessment of whether the variability manipulation influenced the learning of the trained items themselves, or to make it easy to measure how much performance degrades as a function of how far away testing items are from the training items.

The influence of training variation has received a large amount of attention in the domain of sensorimotor skill learning. Much of this research has been influenced by the work of (@schmidtSchemaTheoryDiscrete1975), who proposed a schema-based account of motor learning as an attempt to address the longstanding problem of how novel movements are produced. Schema theory presumes that learners possess general motor programs for a class of movements (e.g. an underhand throw). When called up for use motor programs are parameterized by schema rules which determine how the motor program is parameterized or scaled to the particular demands of the current task. Schema theory predicts that variable training facilitates the formation of more robust schemas, which will result in improved generalization or transfer. Experiments that test this hypothesis are often designed to compare the transfer performance of a constant-trained group against that of a varied-trained group. Both groups train on the same task, but the varied group practices with multiple instances along some task-relevant dimension that remains invariant for the constant group. For example, studies using a projectile throwing task might assign participants to either constant training that practicing throwing from a single location, and a varied group that throws from multiple locations. Following training, both groups are then tested from novel throwing locations [@pigottMotorSchemaStructure1984; @willeyLimitedGeneralizationVaried2018; @pachecoLearningSpecificIndividual2018; @wulfEffectTypePractice1991].

On occasion the Kerr and Booth design may be nested within a larger experimental design. One such study that used a movement timing task, wherein subjects had to move their hand from a starting location, to a target location, attempting to arrive at the target location at specific time following the onset of a cue [@wrisbergVariabilityPracticeHypothesis1987]. This study utilized 4 different constant groups, and 3 varied groups, with one of the constant groups training under conditions identical to the testing conditions, and which were not trained on by any of the varied groups, e.g. the design of Kerr and Booth. However, in this case the varied group did not outperform the constant group.

 A more recent study attempting a slightly more direct replication of the original Kerr & Booth study [@willeyLongtermMotorLearning2018], having subjects throw beanbags at a target, with the varied group training from positions (5 and 9 feet) on either side of the constant group (7 feet). However, this study diverged from the original in that the participants were adults; they faced away from the target and threw the beanbag backwards over their bodies; they alternated using their right and left hands every 6 trials; and underwent a relatively extreme amount of training (20 sessions with 60 practice trials each, spread out over 5-7 weeks). Like @wrisbergVariabilityPracticeHypothesis1987, this study did not find a varied advantage from the constant training position, though the varied group did perform better at distances novel to both groups.

Of course, the relationship between training variability and transfer is unlikely to be a simple function wherein increased variation is always beneficial. Numerous studies have found null, or in some cases negative effects of training variation [@deloshExtrapolationSineQua1997; @sinkeviciuteRoleInputVariability2019; @wrisbergVariabilityPracticeHypothesis1987], and many more have suggested that the benefits of variability may depend on additional factors such as prior task experience, the order of training trials, or the type of transfer being measured [@bernikerEffectsTrainingBreadth2014; @braithwaiteEffectsVariationPrior2015; @hahnEffectsCategoryDiversity2005; @lavanEffectsHighVariability2019; @northEffectConsistentVaried2019; @sadakataIndividualAptitudeMandarin2014; @zamanPerceptualVariabilityImplications2021].

@@

### 2024-04-11 Sections/hold_temp.qmd
intro update

>>>>

## Schmidt

A noteworthy early study that explored the benefits of variability, as postulated by Schmidt, was conducted by Kerr & Booth (1978). In their experiment, two groups of children, ages 8 and 12, were tasked with a bean bag throwing exercise. The constant training group practiced aiming at a target 3 feet away, while the varied training group practiced from distances of both 2 and 4 feet. During this training phase, participants were blindfolded to prevent visual targeting, though they could check where the beanbag landed after each attempt for feedback. Twelve weeks later, all participants were tested at the novel distance of 3 feet, which was familiar to the constant group but new to the varied group. Again blindfolded, they received feedback only after the first half of their throws during the final test. Notably, both age groups exhibited significantly better performance in the varied condition, with a more pronounced effect in the younger cohort. Although this study design indirectly supports the hypothesis that varied training promotes better generalization than constant training, it robustly demonstrates the efficacy of varied practice without directly testing novel positions for the constant group.

## Literature Review

An early and influential work on the influence of variability on category learning is that of Posner & Keele (1968). In an ambitious attempt to address the question of how category information is represented, the authors trained participants to categorize artificial dot patterns, manipulating whether learners were exposed to examples clustered close to the category prototypes (e.g. a low variability condition), or spread further away from the prototype (the varied-training group). It should be noted that both groups in this study were trained with the same number of unique instances and the manipulated difference was how spread out the instances were. The authors claim based on prior experiments using the same stimuli, that the training stimuli for the varied group were at least as far away from the testing stimuli as the training stimuli of the less-varied group. The authors interpreted their findings as evidence for the extraction of an abstraction or schema that is extracted and stored, and then over time becomes more likely to be the reference point from which generalization occurs, given that specific instances are thought to decay at a faster rate than prototypes or schema. The Posner and Keele study has been extremely influential and continues to be cited in contemporary research as clear evidence that schema abstraction underlies the benefits of varied training. It's also referenced as a key influence in the development of "Schema Theory of Motor Learning" (Schmidt, 1975), which in turn influenced decades of investigations on the potential benefits of varied training in motor skill learning. However, the classic Posner & Keele study despite being far more carefully designed than many subsequent studies, and despite being a relative rarity in explicitly discussing and attempting to control for potential confounds of similarity between groups, may nevertheless be emblematic of a common issue in many investigations of the effects of varied training on learning. The problem with Posner & Keele's conclusion was demonstrated clearly almost 3 decades later (Palmeri & Nosofsky, 2001), when researchers conducting a near replication of the original study also collected similarity judgements following training and performed multidimensional scaling analysis. Rather than being in the middle of the training stimuli as was the case in the physical stimuli space, the psychological representation of the prototype was shown reside at an extreme point, and generalization patterns by participants that would have seemed to warrant the learning of a prototype were then easily accounted for with only the assumption that the participants encoded instances. One of the primary concerns of the present paper is that many of the studies which purport to explain the benefits of variation via prototypes, schemas, or other abstractions, are often overlooking the potential of instance based similarity accounts.
